Tight Bounds for Online Balanced Graph Partitioning

### 2020-05-08 paper/disc20clustering.tex
k=3 addressing comments

 \title{Tigth Bounds for Online Balanced Graph Partitioning}

@@

Now  we discuss the general online
model where the request sequence
can be arbitrary.
In Section~\ref{sec:k3}, we show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} provides an optimal algorithm for $k=3$.
Then, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.

@@

\noindent
\textbf{Component-based algorithm.}
Now we describe the algorithm \TAlg that we use throughout this section.
It partitions nodes into components, and
initially, each node belongs to its own singleton component.
For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.3\textwidth]{figs/substitute}
	\caption{Cluster signatures in warm-up k=3 proof}
\end{figure}

In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
Before bounding the competitive ratio of \TAlg, we introduce the lemma that estimates the cost of a single repartition of \TAlg.

@@

	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
	Note that $U\neq V$,
	as otherwise, the request would not trigger a~repartitioning.
	We consider cases based on the types of clusters $U$ and $V$.
	Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, we merge components of size $3$, and no component respecting partitioning exists, a contradiction.
	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
	For two clusters, any configuration can be reached within two swaps, due to the fact that clusters are indistinguishable.

	Finally, we consider the case where both $U$ and $V$ are of type $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists. 
	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists. In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
\end{proof}

>>>>

 \title{Tight Bounds for Online Balanced Graph Partitioning}

@@

Now  we discuss the general online
model where the request sequence
can be arbitrary.
In Section~\ref{sec:k3}, we show that the classic \emph{rent-or-buy} approach~\cite{karlin-ski-rental} provides an optimal algorithm for $k=3$.
We showed a lower bound of $\Omega(k \cdot \ell)$ that holds for $k=3$ (cf. Section~\ref{sec:lowerbound}), hence the result from this section is asymptotically optimal.

Then, in Section~\ref{sec:k2}, we show a strictly $6$-competitive algorithm for $k=2$.
\maciek{Intro: compare to previous 7-competitive algorithm. Note strict competitive ratio vs large additive term. Also, emphasize that the algorithm is different than 7-competitive}

@@

\noindent
\textbf{Component-based algorithm.}
Now we describe the algorithm \TAlg that we use throughout this section.
The algorithm partitions nodes into components, and
initially, each node belongs to its own singleton component.
For each pair of nodes $\set{x,y}$, \TAlg maintains a counter $C_{\set{x,y}}$ and increments it on every external request between $x$ and $y$.
Once $C_{\set{x,y}} = \alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partitioning.
If no such partitioning exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.3\textwidth]{figs/substitute}
	\caption{Cluster types used in the analysis of \TAlg.}
\end{figure}

In our analysis, we distinguish among three types of clusters: $C_1, C_2, C_3$. In a cluster of type $C_i$, the size of the largest component contained in this cluster is $i$.
Before bounding the competitive ratio of \TAlg, we introduce the lemma that upper bounds the cost of a single repartition of \TAlg.

@@

	Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
	The request triggered the repartition, hence it was external and $U\neq V$.
	We consider cases based on the types of clusters $U$ and $V$.

	If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the reconfiguration is local within $U$ and $V$.
	In this case, a local repartition is possible (within $U$ and $V$), for a cost of at most $2$ swaps.
	Note that it is impossible that either $U$ or $V$ is of type $C_3$, as otherwise, a components of size $3$ participates in a merge, which results in a component of size at least $4$, and in this case no component respecting partitioning exists, a contradiction.

	It remains to consider the case where both $U$ and $V$ are of type $C_2$. Note that $(u,v)$ cannot both belong to a component of size $2$, as this would mean that \TAlg has the component of size $4$, a contradiction with the case assumption that the component respecting partitioning exists. 
	Otherwise, if either $u$ or $v$ belongs to a component of size $2$, then it suffices to exchange components of size $1$ between $U$ and $V$.
	Finally, if $u$ and $v$ belong to components of size $1$, then we must place them in a cluster different from $U$ and $V$.
	Note that in such case, a $C_1$-type cluster $W$ exists, as otherwise no component respecting partitioning exists. In this case \TAlg performs one swap, i.e., it exchanges the nodes $u$ and $v$ with any two nodes of $W$.
\end{proof}
